---
title: "TP: ggplot"
author: "Eric Marcon"
date: "`r format(Sys.time(), '%d %B %Y')`"
url: https://EricMarcon.github.io/Cours-R-Geeft/
github-repo: EricMarcon/Cours-R-Geeft
bibliography: references.bib
biblio-style: chicago
urlcolor: blue
output:
  bookdown::beamer_presentation2:
    latex_engine: xelatex
    includes:
      in_header: latex/header.tex
    citation_package: natbib
    slide_level: 2
    df_print: default
    number_sections: no
    toc: no
    fig_caption: no
    keep_tex: no
  bookdown::ioslides_presentation2:
    logo: images/logo.png
    widescreen: true
---

```{r}
#| label: DoNotModify
#| include: false
### Utilities. Do not modify.
# Installation of packages if necessary
InstallPackages <- function(Packages) {
  InstallPackage <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {
      install.packages(Package, repos = "https://cran.rstudio.com/")
    }
  }
  invisible(sapply(Packages, InstallPackage))
}

# Basic packages
InstallPackages(c("bookdown", "formatR", "kableExtra", "ragg"))

# kableExtra must be loaded 
if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx") {
  # Word output (https://stackoverflow.com/questions/35144130/in-knitr-how-can-i-test-for-if-the-output-will-be-pdf-or-word)
  # Do not use autoformat (https://github.com/haozhu233/kableExtra/issues/308)
  options(kableExtra.auto_format = FALSE)
}
library("kableExtra")

# Chunk font size hook: allows size='small' or any valid Latex font size in chunk options
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r}
#| label: Options
#| include: false
### Customized options for this document
# Add necessary packages here
Packages <- c(
  "tidyverse", 
  "sf"
)
# Install them
InstallPackages(Packages)

# knitr options
knitr::opts_chunk$set(
  cache =   TRUE,     # Cache chunk results
  include = TRUE,     # Show/Hide chunks
  echo =    TRUE,     # Show/Hide code
  warning = FALSE,    # Show/Hide warnings
  message = FALSE,    # Show/Hide messages
  # Figure alignment and size
  fig.align = 'center', out.width = '80%',
  # Graphic devices (ragg_png is better than standard png)
  dev = c("ragg_png", "pdf"),
  # Code chunk format
  tidy = FALSE, tidy.opts = list(blank = FALSE, width.cutoff = 50),
  size = "scriptsize", knitr.graphics.auto_pdf = TRUE
)
options(width = 50)

# ggplot style
library("tidyverse")
theme_set(theme_bw())
theme_update(
  panel.background = element_rect(fill = "transparent", colour = NA),
  plot.background = element_rect(fill = "transparent", colour = NA)
)
knitr::opts_chunk$set(dev.args = list(bg = "transparent"))

# Tibbles: 5 lines, fit to slide width
options(tibble.print_min = 5, tibble.width = 50)

# Random seed
set.seed(973)
```


# Ventoux

## Données

```{r}
read_csv2("data/Inv_GEEFT_Ventoux_09-2020.csv") |> 
  rename(
    espece = Espèce, 
    diametre = `Diamètre (cm)`, 
    hauteur = `Hauteur réelle (m)`
  ) |> 
  mutate(
    espece = case_match(
      espece, 
      "P" ~ "Pin",
      "C" ~ "Cèdre"
    )
  ) ->
  ventoux
```


## Graphique hauteur ~ diamètre

```{r}
#| out.width: 70%
ventoux |> 
  ggplot(aes(x= diametre, y = hauteur)) +
  geom_point(aes(col = espece)) +
  geom_smooth(method = "lm")
```


## Esthétique

Comparer avec 

```{r}
#| out.width: 70%
ventoux |> 
  ggplot(aes(x= diametre, y = hauteur, col = espece)) +
  geom_point() +
  geom_smooth(method = "lm")
```


# Paracou

## Données

```{r}
read_csv2("data/Paracou6.csv") |> 
  unite(
    # Nouvelle colonne
    col = spName, 
    # Champs concaténés
    Genus, Species, 
    # Séparateur
    sep = " ",
    # Conserver les colonnes d'origine
    remove = FALSE
  ) ->
  paracou6
  
```

$\to$ Examiner paracou6


## Carte des Wapa

Objectif : cartographier les Wapas (genre : Eperua) par espèces

- Filter le genre Eperua ;
- Ajouter une colonne pour la surface terrière ;
- Sélectionner les colonnes contenant le nom de l'espèce, la position et la surface terrière


## Bagarre

```{r}
paracou6 |> 
  filter(Genus == "Eperua") |> 
  mutate(G = (CircCorr / 2 / pi)^2 * pi / 10000 ) |> 
  select(Xfield, Yfield, spName, G) |> 
  print() ->
  paracou6_wapa
```

## Carte

```{r}
library("sf")
paracou6_wapa |> 
  st_as_sf(coords = c("Xfield", "Yfield")) |> 
  ggplot() +
  geom_sf(
    aes(col = spName, size = G)
  ) +
  labs(col = "Espèce")
```



<!-- Styles for HTML slides -->
<style>
  /* Allow long bibliography */
  .forceBreak { -webkit-column-break-after: always; break-after: column; }
  slides > slide { overflow: scroll; }
  slides > slide:not(.nobackground):after { content: ''; }

  /* First page logo size */
  .gdbar img {
    width: 200px !important;
    height: 55px !important;
    margin: 8px 8px;
  }
  .gdbar {
    width: 250px !important;
    height: 70px !important;
  }
  
  /* No logo on slides */
  slides > slide:not(.nobackground):before {
    display:none
  }
</style>
